# Getting and Cleaning the data: Peer Assignment 2

For the following analysis i have used R 3.2.2 and as OS Windows 10.
I used comments before the download-commands to prevent a download in each execution.

## Download and importing the raw data
Firstly we set the path, download the zip-archive and load the data into tabels.

```{r echo=TRUE}
path <- getwd()
#download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip", "run.zip" )
#unzip("run.zip",list=T)
path = paste(path,"/UCI HAR Dataset/",sep="")
setwd(path)

train_fea <- read.table("train/X_train.txt",sep="",header=F)
train_subject <- read.table("train/subject_train.txt",header = F)
train_activity <- read.table("train/y_train.txt",header=F)

test_fea <- read.table("test/X_test.txt",sep="",header=F)
test_subject <- read.table("test/subject_test.txt",header=F)
test_activity <- read.table("test/y_test.txt",header=F)
```

After that we read the name of the features from the according text-file and give all variables the correct name. An extra variable is added called "trainortest". It is a factor tracking if the record is part of the training or the test set.

```{r}
setwd(path)
names_var <- read.table("features.txt",sep="")
train <- cbind("train",train_subject,train_activity,train_fea)
test <- cbind("test",test_subject,test_activity,test_fea)

colnames(train) <- c("trainortest","subject","activity",as.character(names_var$V2))
colnames(test) <- c("trainortest","subject","activity",as.character(names_var$V2))
```

## 1. Merge the training and the test sets to create one data set.

Now the training and test sets are merged together.

```{r}
whole <- rbind(train,test)
whole$subject <- as.factor(whole$subject)
whole$activity <- as.factor(whole$activity)
```

## 2. Extract only the measurements on the mean and standard deviation for each measurement

Now we want to focus only on the means and standard deviations of the measurements. Therefore we search for those variables which have "mean()" or "std()" in their name.

```{r}
index <- grep("mean()",colnames(whole)[4:length(colnames(whole))], fixed = TRUE)
index2 <- c(index,grep("std()",colnames(whole)[4:length(colnames(whole))], fixed =TRUE))
whole_stdmean <- whole[,c(1:3,(index2+3))] 
```

## 3. Use descriptive activity names to name the activities in the data set

Instead of using integer numbers noting the activity, we load the according activity labels and rename the factor values.

```{r}
setwd(path)
activity_labels <- read.table("activity_labels.txt")
levels(whole_stdmean$activity)[1:6]<- as.character(activity_labels$V2)
```


## 4. Appropriately label the data set with descriptive variable names

Now we want to give the variables more descriptive names.

```{r}
names_variables <- colnames(whole_stdmean)[-(1:3)]
new_names <- gsub("^t","Time",names_variables)
new_names2 <- gsub("Acc","Acceleration",new_names)
new_names3 <- gsub("Gyro","Gyroscope",new_names2)
new_names4 <- gsub("Mag","Magnitude",new_names3)
new_names5 <- gsub("^f","Frequency",new_names4)
new_names6 <- gsub("mean()","in_average", new_names5, fixed=TRUE)
new_names7 <- gsub("std()","standard-deviation", new_names6, fixed=TRUE)
new_names7 <- gsub("X","X-axis", new_names7)
new_names7 <- gsub("Y","Y-axis", new_names7)
new_names7 <- gsub("Z","Z-axis", new_names7)

colnames(whole_stdmean)[-(1:3)]<- new_names7
```


## 5. Creates a second, independent tidy data set with the average of each variable for each activity and each subject

As a second, tidy data set we want a dataframe which has one record for each existing combination of subject-id and activity. As additional variables we compute the average value of each non-excluded measurement over the according subset. 

```{r}
averages <- numeric(180*66)
trainortest <- factor(character(180),levels=levels(whole_stdmean$trainortest))
subject_avg <- factor(character(180),levels=levels(whole_stdmean$subject))
activity_avg <- factor(character(180),levels=levels(whole_stdmean$activity))


dim(averages) <- c(180,66)
i <- 0
for (subj in levels(whole_stdmean$subject)){ 
    for (acti in levels(whole_stdmean$activity)){
        i <- i+1
        wholesubset <- whole_stdmean[whole_stdmean$subject==subj & whole_stdmean$activity==acti,]
        avg <- colMeans(wholesubset[,(4:length(wholesubset[1,]))])
        averages[i,] <- avg
        trainortest[i] <- wholesubset$trainortest[1]
        subject_avg[i] <- subj
        activity_avg[i] <- acti
    } 
}
df_avg <- data.frame(trainortest,subject_avg,activity_avg,averages)
colnames(df_avg)<-colnames(whole_stdmean)    
```

There are two tidy data sets now: whole_stdmean and df_avg.
That´s my analysis, i hope you appreciate it! Thanks for reviewing.